An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-03 nips_2016.tex
Slight clarification to appendix b

In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
parameter is a 2-D vector
$\theta = \langle \theta_1, \theta_2 \rangle$, where
\begin{equation}
\theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
\end{equation}
and where $\mathcal{N}$ indicates the normal distribution (more generally, the
multivariate normal). We consider the above as our prior.
Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
log prior probability we endow on $\theta$ is
\begin{equation}
\log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
\end{equation}
To generate the data, we use the following Gaussian mixture with tied means:
\begin{equation}\label{eq:x_points}
x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
\end{equation}
where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means
the log likelihood of a single data instance is
\begin{equation}
\log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
\log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
(\theta_1+\theta_2))^2\right)\right)
\end{equation}
The official problem statement is as follows: given some number of conditionally
independent data points $x_1, x_2, \ldots, x_N$ generated according
to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
\begin{equation}\label{eq:log_post}
\log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
\end{equation}
Alternatively, depending on the application, we may opt to instead pick a point
estimate\footnote{We do this for the logistic regression experiment, described
in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
is extremely large, it will cause the posterior to peak sharply at its modes,
reducing distribution estimates to point estimates.)

>>>>

In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
parameter is a 2-D vector
$\theta = \langle \theta_1, \theta_2 \rangle$, where
\begin{equation}
    \theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
\end{equation}
and where $\mathcal{N}$ indicates the normal distribution (more generally, the
multivariate normal). We consider the above as our prior.
Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
log prior probability we endow on $\theta$ is
\begin{equation}
    \log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
\end{equation}
To generate the data, we use the following Gaussian mixture with tied means:
\begin{equation}\label{eq:x_points}
    x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
\end{equation}
where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
arbitrary, and different choices would yield different posterior shapes).
This means the log likelihood of a single data instance is
\begin{equation}\label{eq:log_ll_single_instance}
    \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
    \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
    (\theta_1+\theta_2))^2\right)\right)
\end{equation}
The official problem statement is as follows: given some number of conditionally
independent data points $x_1, x_2, \ldots, x_N$ generated according
to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
\begin{equation}\label{eq:log_post}
    \log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
\end{equation}
Alternatively, depending on the application, we may opt to instead pick a point
estimate\footnote{We do this for the logistic regression experiment, described
in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
is extremely large, it will cause the posterior to peak sharply at its modes,
reducing distribution estimates to point estimates.)
